---
title: "MIS 381N Homework 1"
author: "Sam Malcolm (stm2334)"
output:
  pdf_document: default
---
# Problem 1

We know that $(AB)^{-1} = B^{-1}A^{-1}$ so we also know that $(AB)^T = B^TA^T$.

We also know that the multiplicative identity $I$ aka $AA^{-1}$ is symmetric, so $I^T = I$.

By setting $B = A^T$ and multiplying to each side, we see:

\begin{align}
(A^{-1})^{T}\times A^{T} &= (A^{T})^{-1}\times A^{T}\\
(AA^{-1})^{T} &= I\\
I^{T} &= I\\
I &= I\\
\end{align}

This proves that $(A^{-1})^{T} = (A^{T})^{-1}$.

# Problem 2

There are four different types of loans - First mortgage, Second mortgage, Home improvement, Personal overdraft - which are represented respectively by $x_{1}$,$x_{2}$,$x_{3}$,and $x_{4}$ . 

Based on the given information, we know that:
\begin{equation}
  \begin{cases}
  x_{1} + x_{2} + x_{3} + x_{4} &= 250,000,000\\
  0.45x_{1} - 0.55x_{2}&=0\\
  x_{2} &= 62,500,000\\
  0.14x_{1} + 0.2x_{2} + 0.2x_{3} + 0.1x_{4} &= 150,000,000\\
  \end{cases}
\end{equation}

We can use this to create three matrices:

\begin{equation}
A = 
  \begin{bmatrix}
  1 & 1 & 1 & 1 \\
  .45 & -.55 & 0 & 0 \\
  0 & 1 & 0 & 0 \\
  0.14 & 0.2 & 0.2 & 0.1 \\
  \end{bmatrix},
\quad
\textbf{x} = 
  \begin{bmatrix}
  x_{1}\\
  x_{2}\\
  x_{3}\\
  x_{4}\\
  \end{bmatrix},
\quad
\textbf{y} = 
  \begin{bmatrix}
  250,000,000\\
  0\\
  62,500,000\\
  37,500,000\\
  \end{bmatrix}
\end{equation}

We can then solve for the four loan amounts by setting $A\textbf{x} = \textbf{y}$ aka $\textbf{x} = A^{-1}\textbf{y}$.
```{r}
y = c(250000000,0,62500000,37500000)
A = matrix(c(1,.45,0,.14,1,-.55,1,.2,1,0,0,.2,1,0,0,.1),4,4) 
x = solve(A,y)
x_mat = matrix(c(x),4,1)
print(x_mat)
```

# Problem 3

We want to choose  $x_{1}$,$x_{2}$,$x_{3}$,$x_{4}$

to maximize

\[
1.5x_{1} + 2.5x_{2} + 3x_{3} + 4.5x_{4}
\]

subject to the constraints

\begin{equation}
  \begin{cases}
  2x_{1} + 4x_{2} + 3x_{3} + 7x_{4} &\leq 100000\\
  3x_{1} + 2x_{2} + 3x_{3} + 4x_{4} &\leq 50000\\
  2x_{1} + 3x_{2} + 2x_{3} + 5x_{4} &\leq 60000\\
  x_{1},x_{2},x_{3},x_{4} &\geq 0\\
  \end{cases}
\end{equation}

In matrix notation, we can set this up as:

\begin{equation}
\textbf{x} = 
  \begin{bmatrix}
  x_{1}\\
  x_{2}\\
  x_{3}\\
  x_{4}\\
  \end{bmatrix},
\quad
\textbf{C} = 
  \begin{bmatrix}
  1.5\\
  2.5\\
  3\\
  4.5\\
  \end{bmatrix},
\quad
A = 
  \begin{bmatrix}
  2 & 4 & 3 & 7 \\
  3 & 2 & 3 & 4 \\
  2 & 3 & 2 & 5 \\
  \end{bmatrix},
\quad
\textbf{b} = 
  \begin{bmatrix}
  100,000\\
  50,000\\
  60,000\\
  \end{bmatrix}
\end{equation}

We can then maximize $\textbf{C}^{T}\textbf{x}$ subject to $A\textbf{x} \leq \textbf{b}$ and ${x} \geq 0$

# Problem 4

## Part a

```{r}
A = matrix(nrow=20, ncol=20)
A[1,1] = 1
for(j in 1:20){
   for(i in 1:20){
     if (j > i){
        A[i,j] = i/j
      }
     else if (j == i){
       A[i,j] = i/j
      }
     else{
       A[i,j] = j/i
     }
   }
}
A
```

## Part b

```{r}
all.equal(t(A),A)
```

## Part c

```{r}
C = solve(A)
C
```
```{r}
ident = diag(20)
product = A%*%C
all.equal(product, ident)
```

## Part d

```{r}
d = c(1,2,3,4,5,6,7,8,9,10,10,9,8,7,6,5,4,3,2,1)
d
```

## Part e

```{r}
x = solve(A, C%*%d)
x
```

